Software Architectures to Integrate Workflow Engines in Science Gateways

### 2016-04-27 paper.tex
Fixed note in conclusion

The pool architecture is overall the most scalable, which is not a
surprise since it was designed precisely for scalability. Tight
integration is the least scalable and all the other architectures
perform the same. The overall scalability score, however, should not
mask the unique characteristics of architectures regarding this
criterion. Nested workflows are the only architecture that can easily
accommodate distributed workflow execution, which can be critical in
some cases. At the same time, the scheduling constraints created by
the sub-tasking and nested workflow architectures could well become
showstoppers depending on the type of the targeted
infrastructure. Non-reliable infrastructures, for example, could
hardly cope with workflow engines being wrapped in computing tasks as
done in sub-tasking \note{Marc-e}{free for all? WE dev involved?}. The availability of multiple engine instances
could also become a critical feature for science gateways with
important workloads, which would favor sub-tasking and pool.

@@

% Summary of contributions
We reviewed architectures used to integrate workflow engines in
science gateways, we described them in a system-independent framework
suitable for comparison, we illustrated them on real systems, we
developed quantitative metrics for evaluation, we evaluated these
metrics for all the architectures, and we discussed the pros and cons
of all the architectures based on these metrics.
\closedanswerednote{Rafael}{One thing that is missing in the
  conclusion or the limitations section is to state that this work
  does not intend to quantify the amount of work required to develop
  science gateways using a specific architecture, but the goal is to
  highlight what are the main components and interactions for each
  architecture, and the main advantages and drawbacks. This should
  also be stated in the introduction.}{Tristan}{This is in
  contradiction with the metric on 'integration effort'. What do you
  think?}{Tristan}{Fixed, as discussed in github \#8.}

% Impact of the work
To the best of our knowledge, our work is the first of its kind
\note{Tristan}{Be more accurate}. Our results can be used as
recommendation for science gateway and workflow engine architects to
build new platforms, or to upgrade existing ones.

@@

This work is in the scope of the LABEX PRIMES (ANR-11- LABX-0063) of
Universit\'e de Lyon, within the program ``Investissements d’Avenir''
(ANR-11-IDEX-0007) operated by the French National Research Agency
(ANR). This work also falls into the scope of the scientific topics of
the French National Grid Institute (IdG). We thank the site
administrators of the European Grid Initiative and the GGUS support
for their help related to the VIP platform. The CBRAIN team is grateful for the
computing cycles, storage, and support obtained from Compute Canada
(\url{https://computecanada.ca}) and platform development program from CANARIE
 (\url{http://www.canarie.ca/}). This work has been made possible with the support of the Irving Ludmer Family Foundation and
the Ludmer Centre for Neuroinformatics and Mental Health. \answerednote{Marc-e}{Would it be proper to thank EGI?
  Compute Canada? CANARIE?}{Tristan}{Added EGI and Compute Canada. Not sure if
  Canarie should be added. What about Calcul QC?}{Marc-e}{CQ, no. CANARIE, I'd
  say yes since they funded a major chunk of CBRAIN's development.}

>>>>

The pool architecture is overall the most scalable, which is not a
surprise since it was designed precisely for scalability. Tight
integration is the least scalable and all the other architectures
perform the same. The overall scalability score, however, should not
mask the unique characteristics of architectures regarding this
criterion. Nested workflows are the only architecture that can easily
accommodate distributed workflow execution, which can be critical in
some cases. At the same time, the scheduling constraints created by
the sub-tasking and nested workflow architectures could well become
showstoppers depending on the type of the targeted
infrastructure. Non-reliable infrastructures, for example, could
hardly cope with workflow engines being wrapped in computing tasks as
done in sub-tasking. The availability of multiple engine instances
could also become a critical feature for science gateways with
important workloads, which would favor sub-tasking and pool.

@@

% Summary of contributions
We reviewed architectures used to integrate workflow engines in
science gateways, we described them in a system-independent framework
suitable for comparison, we illustrated them on real systems, we
developed quantitative metrics for evaluation, and we discussed the
pros and cons of all the presented architectures based on these metrics.
\closedanswerednote{Rafael}{One thing that is missing in the
  conclusion or the limitations section is to state that this work
  does not intend to quantify the amount of work required to develop
  science gateways using a specific architecture, but the goal is to
  highlight what are the main components and interactions for each
  architecture, and the main advantages and drawbacks. This should
  also be stated in the introduction.}{Tristan}{This is in
  contradiction with the metric on 'integration effort'. What do you
  think?}{Tristan}{Fixed, as discussed in github \#8.}

% Impact of the work
To the best of our knowledge, our work is the first of its kind. So
far, the literature on science gateways and workflow engines has
focused on the description of particular systems, or on the
presentation of a particular architecture. For instance, the work
in~\cite{Gannon2005} discusses service-oriented architectures for
science gateways with workflow engines.  Instead, our analysis
abstracts and evaluates architectural patterns independently from any
particular system. It provides general insight to integrate science
gateways and workflow engines.

@@

This work is in the scope of the LABEX PRIMES (ANR-11- LABX-0063) of
Universit\'e de Lyon, within the program ``Investissements d’Avenir''
(ANR-11-IDEX-0007) operated by the French National Research Agency
(ANR). This work also falls into the scope of the scientific topics of
the French National Grid Institute (IdG). We thank the site
administrators of the European Grid Initiative and the GGUS support
for their help related to the VIP platform. The CBRAIN team is grateful for the
computing cycles, storage, and support obtained from Compute Canada
(\url{https://computecanada.ca}) and platform development program from CANARIE
 (\url{http://www.canarie.ca/}). This work has been made possible with the support of the Irving Ludmer Family Foundation and
the Ludmer Centre for Neuroinformatics and Mental Health. \closedanswerednote{Marc-e}{Would it be proper to thank EGI?
  Compute Canada? CANARIE?}{Tristan}{Added EGI and Compute Canada. Not sure if
  Canarie should be added. What about Calcul QC?}{Marc-e}{CQ, no. CANARIE, I'd
  say yes since they funded a major chunk of CBRAIN's development.}
